The 4th Workshop on Sustainable Software: Best Practices and Experiences (WSSSPE4)

### 2016-06-10 WSSSPE4_proposal/WSSSPE4_proposal.tex
change in summary and intro

{\bf \noindent Intellectual Merit:} The participants will be selected using a formal application procedure, with selection criteria based around (1) the potential of their attendance to contribute towards their own research and educational goals; (2) their potential contribution to the workshop; (3) providing a diverse set of attendees with priority to early stage researchers (e.g. students, postdoctoral researchers, early stage faculty). The workshop will include presentations, panels and discussions driven from short peer reviewed papers that will advance understanding in this new area of attention.

{\bf \noindent Broader Impacts:} All material from the workshop will be openly archived on public popular sites (e.g. Figshare, Slideshare, arXiv), it is anticipated that the workshop will again be tweeted and documented live including contributions from external participants. It is expected that the workshop will lead to a special issue of an open journal. The organizers take seriously the responsibility to encourage participation from underrepresented groups as workshop leaders as well as attendees, and NSF funds will be used to encourage such participation from the US community.

@@

The first WSSSPE workshop (see Appendix for a description of WSSSPE 1) was held as part of Supercomputing 2013 in Denver during November 2013. The aim on the workshop was to bring together researchers and practitioners to discuss challenges and issues around the development, deployment and maintenance of reusable software for academic research and education. The next workshop, WSSSPE 2, will continue these discussions, focusing on some of the themes identified during WSSSPE~1. WSSSPE~2 is anticipated to be held as a workshop as part of the International IEEE Supercomputing 2014 conference in New Orleans in November 2014. A proposal has been submitted to SC14 which will be reviewed in the next month.

The goals of WSSSPE 2 are
\begin{itemize}
\item Provide an effective mechanism to further discussions and identify research topics in the development, deployment and maintenance of reusable software
\item Proactively encourage the inclusion of issues related to education and involvement of industry
\item Develop a diverse community of engaged researchers, software developers, industrial representatives, students and educators
\item Make the event as accessible and inclusive as possible, leveraging social media and other technologies
\item Motivate the community in providing outputs including publications and reports
\end{itemize}

>>>>

{\bf \noindent Intellectual Merit:} The participants will be selected using a formal application procedure, with selection criteria based around (1) the potential of their attendance to contribute towards their own research and educational goals; (2) their potential contribution to the workshop; (3) providing a diverse set of attendees with priority to early stage researchers (e.g. students, postdoctoral researchers, early stage faculty). The workshop will include vision and position papers as well as lightning talks and demos, aimed at improving sustainable scientific software today as well as defining and working towards the its future.

{\bf \noindent Broader Impacts:} WSSSPE4 accepted submissions (except presentations of previously published work) are planned to be published by WSSSPE as a special collection in http://ceur-ws.org/, which is fully indexed.
Other material from the workshop will be openly archived on public popular sites (e.g. Figshare, Slideshare, arXiv). It is anticipated that the workshop will again be tweeted and documented live including contributions from external participants. The organizers take seriously the responsibility to encourage participation from underrepresented groups as workshop leaders as well as attendees, and NSF funds will be used to encourage such participation from the US community.

@@

Progress in scientific research is dependent on the quality and accessibility of research software at all levels. It is now critical to address many new challenges related to the development, deployment, maintenance, and sustainability of open-use research software: the software upon which specific research results rely.  Open-use software means that the software is widely accessible (whether open source, shareware, or commercial).  Research software means that the choice of software is essential to specific research results; using different software could produce different results.
In addition, it is essential that scientists, researchers, and students are able to learn and adopt a new set of software-related skills and methodologies. Established researchers are already acquiring some of these skills, and in particular, a specialized class of software developers is emerging in academic environments who are an integral and embedded part of successful research teams. WSSSPE provides a forum for discussion of these challenges, including both positions and experiences, and a forum for the community to assemble and act.

This is the fourth major event in the WSSSPE series. The first two WSSSPE workshops were held as part of SC in 2013 and 2014. These one-day workshops brought together researchers and practitioners to discuss challenges and issues around the development, deployment and maintenance of reusable software for academic research and education, and discussed potential solutions, moving from a presentation to a presentation and discussion format.  WSSSPE3, held at NCAR in 2015, grew out of the previous events, and focused almost entirely on facilitated discussions, with the aim of starting working groups to make changes in the practice of scientific software development leading to more sustainable software.  Two of these groups have continued
working together, leading to the FORCE11 Software Citation Principles and an in-progress paper
on best practices for sustainable scientific software development.

In WSSSPE4, held over 2 1/2 days in Manchester, UK, we will include elements from the
previous workshops, with most of the workshop divided into two tracks:

{\bf Track 1 – Building a sustainable future for open-use research software} has the goals of defining a vision of the future of open-use research software, and in the workshop, initiating the activities that are needed to get there.  The idea of this track is to first think about where we want to be 5 to 10 years from now, without being too concerned with where we are today, and then to determine how we can move to this future.

{\bf Track 2 – Practices \& experiences in sustainable scientific software} has the goal of improving the quality of today’s research software and the experiences of its developers by sharing practices and experiences.  This track is focused on the current state of scientific software and what we can do to improve it in the short term, starting with where we are today.

Topics of interest include but are not limited to:
\begin{itemize}
\item Development and Community
\begin{itemize}
\item Best practices for developing sustainable software
\item Models for funding specialist expertise in software collaborations
\item Software tools that aid sustainability
\item Academia/industry interaction
\item Refactoring/improving legacy scientific software
\item Engineering design for sustainable software
\item Metrics for the success of scientific software
\item Adaptation of mainstream software practices for scientific software
\end{itemize}
\item Professionalization
\begin{itemize}
\item Career paths
\item RSE as a brand
\item RSE outside of the UK or Europe
\item Increase incentives in publishing, funding and promotion for better software
\end{itemize}
\item Training
\begin{itemize}
\item Training for developing sustainable software
\item Curriculum for software sustainability
\end{itemize}
\item Credit
\begin{itemize}
\item Making the existing credit and citation ecosystem work better for software
\item Future credit and citation ecosystem
\item Software contributions as a part of tenure review
\item Case studies of receiving credit for software contributions
\item Awards and recognition that encourage sustainable software
\end{itemize}
\item Software publishing
\begin{itemize}
\item Journals and alternative venues for publishing software
\item Review processes for published software
\end{itemize}
\item Software discoverability/reuse
\begin{itemize}
\item Proposals and case studies
\end{itemize}
\item Reproducibility and testing
\begin{itemize}
\item Reproducibility in conferences and journals
\item Best practices for code testing and code review
\end{itemize}
\end{itemize}

Joint sessions of the workshop will include a keynote, a set of lightning talks on topics related to either track, and a panel on best practices for scientific software development.  Tracks 1 and 2 will be held in parallel with each other.

Track 1 will consist of facilitated discussions, both as a single group and as breakouts.  It will also include discussion of Idea papers that presents implementable proposals related to the track.
Track 1 participants will build on the proposals and ideas in these papers, with the goal of initiating the planning, development, and execution of some of the ideas during the workshop itself. Given the magnitude and importance of the task at hand, the WSSSPE4 organizing committee encourages these proposals to be developed on an open, public, and inclusive basis. Submitters are invited to present a vision of some aspect of the future of open-use research software, and a plan of activities to gather and organize the resources needed to get there.

\begin{quote}
Example idea paper topics include:
\begin{itemize}
\item Adaptation of industrial software engineering principles into the research software community with a plan to fund the work
\item Funding and scaling software carpentry style training in advanced topics
Infrastructure and funding for community maintenance of open use research software
\item Scaling the SSI beyond the UK
\item Specific proposals of how to bridging/network the various research software engineering communities in scalable manner without destroying independence and unique foci of each community
\end{itemize}
\end{quote}

Track 2 will consist mostly of presentations, with some discussions, including position papers, experience papers, presentations of previously published work, and demos.  The intent of this
track is for current scientific software developers and those interested in scientific software development to learn from each other improve the current state of the practice in this area.
